\LARGE \bf Cascading Situation Assessment for Robots

### 2017-02-27 paper.tex
More work on intro

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight possible, user-written
    \emph{clients} (in blue) are exchanging environment models through four shared
    \emph{worlds} (made of joint spatial and temporal models). This architecture
    enables successive and modular refinement of the models
    (\emph{cascading} situation assessment), effectively adapted to each
    client's needs (yellow nodes represent other possible components in the system that do not
    interact directly with the \uwds network).}

@@

However, {\sc Spark} is a monolithic module, that does not aim at sharing its
internal 3D model to other external components. On the contrary, \uwds focuses
on offering a shared and distributed representation within the robot's software
architecture. Previous work on distributed scene graphs~\cite{naef2003blue} has
been applied to robotics to provide a shared 3D representation of the robot's
environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
\emph{Deep State Representation}~\cite{bustos2016unified}).

@@

These different components typically need different models, at different spatial
and temporal resolutions. A 3D motion planner would typically use coarse 3D
models of surrounding objects to lower the computational load while planning,
while a spatial reasoner might need on the contrary a high-resolution model of
the environment for accurate rendering when attempting to estimate the
visibility of an object. Spatial models are also usually large as they might
contain a lot of geometric data (including point clouds and meshes). Efficient
sharing of such datasets is non-trivial, particularly when the said data are
dynamically acquired (\eg in the case of map building or object recognition).

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these complexities: geometric data can be represented, but are not
first-class citizens (basic tasks like displaying a 3D mesh at a dynamic
position are non-trivial with ROS, for instance) and representing and reasoning
on \emph{alternative states} of the environment is not directly feasible.

Representing \emph{alternative states} is actually often critical. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
at the sensor level often introduce hard-to-avoid physical inconsistencies (like
detected objects floating in the air, or on the contrary wrongly inset into
other objects). Therefore, a post-process stage (for instance, using a physics
simulation engine) might be desired to move the objects seen
by the robot into physically-correct positions. Implemented with a classical
approach (for instance, using ROS TF frames), we would get an object {\tt book}
located at the frame {\tt book\_frame}, and a secondary frame {\tt
book\_frame\_corrected} computed by the physics engine. It is easy to see why
such an approach would be quickly confusing (where is in fact my book?) and would not
scale well, with the robot's 3D model cluttered with spurious frames.

@@

serving a scene from
different viewpoints to compute visibility is not directly possible

Relation to Beetz's \emph{deep representation}~\cite{beetz2010towards}

GSM (for \emph{Grounded Situation Model})~\cite{Mavridis2006} is a knowledge
representation system primarily built to ``facilitate cross-modal
interoperability'',  especially in the context of verbal interaction with a
robot.

GSM does not rely on any formal language but rather on a layered data structure
(figure~\ref{fig|gsm}) that organises the surrounding world into agents and
relations between agents.  Each agent (any animate or inanimate object) is
attached to a physical model (made of \emph{body parts} that have properties
like their position, color, etc.) and a mental model (which is a recursively
embedded GSM, thus allowing a sort of theory of mind).

Properties are represented in three layers: a stochastic representation, close
to sensory percepts, a \emph{continuous single-valued} encoding of the
stochastic model, and a discrete, categorical model.

One notable feature of GSM is the \emph{bidirectionality} of the grounding
process: not only sensor percepts are abstracted into categories suitable for
human conversation, but human utterance (like ``There is a red ball in the
center of the table'') can also be turned into property descriptions. This
basically enable the knowledge representation system of the robot to
\emph{imagine} entities.

GSM also features several strategies for managing time and events.
\emph{Moments} are created by storing timestamped snap-shots of GSM, and
\emph{event classifiers} allow to define and detect events.

>>>>

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{overview}
    \caption{Schema of a possible \uwds network: eight possible, user-written
    \emph{clients} (in blue) are sharing environment models through four
    independent \emph{worlds} (made of joint spatial and temporal models). This architecture
    enables successive and modular refinement of the models
    (\emph{cascading} situation assessment), effectively adapted to each
    client's needs (yellow nodes represent other possible components in the system that do not
    interact directly with the \uwds network).}

@@

However, {\sc Spark} is a monolithic module, that does not support sharing its
internal 3D model to other external components. On the contrary, \uwds focuses
on offering a shared and distributed representation within the robot's software
architecture. Previous work on distributed scene graphs~\cite{naef2003blue} has
been applied to robotics to provide a shared 3D representation of the robot's
environment (like the \emph{Robot Scene Graph}~\cite{blumenthal2013scene} or 
\emph{Deep State Representation}~\cite{bustos2016unified} approaches).

@@

These different components typically exhibit different needs in terms of model,
like different nominal spatial and/or temporal resolutions. For instance, a 3D
motion planner would typically use coarse 3D models of surrounding objects to
lower the computational load while planning, while a module assessing the
visibility of objects might need on the contrary a high-resolution model of the
environment for accurate 3D rendering. This requirement of multiple
task-specific representations has been framed as the need for \emph{deep
representation} by Beetz~\cite{beetz2010towards}.

%Spatial models are also usually large as they might
%contain a lot of geometric data (including point clouds and meshes). Efficient
%real-time sharing of such datasets is non-trivial, particularly when the said
%data is changing (\eg in the case of map building or object recognition).

Traditional robotic middlewares like ROS are not particularly well-suited to
deal with these different needs: full geometric data can be represented, but are not
first-class citizens (for instance, basic tasks like displaying a 3D mesh at a dynamic
position are not particularly easy with ROS) and critically, representing and reasoning
on alternative models of the environment is not directly feasible.

Representing \emph{alternative states} is however often highly desirable. For instance,
software components manipulating environment models typically perform better if
the models are physically consistent. However, low-level perception inaccuracies
often introduce hard-to-avoid physical inconsistencies (like
detected objects floating in the air, or on the contrary wrongly inset into
other objects). Therefore, a post-process stage (for instance, using a physics
simulation engine) might be needed to displace the objects seen
by the robot into physically-correct positions. Implemented with a classical
approach (for instance, using ROS TF frames), we would get an object {\tt book}
located at the frame {\tt book\_frame}, and a secondary frame {\tt
book\_frame\_corrected} computed by the physics engine. It is easy to see why
such an approach would be quickly confusing (where is in fact my book?) and would not
scale well, with the robot's 3D model cluttered with spurious frames.

@@

A third example relates to human-robot interaction scenarios where \emph{perspective
taking} is important (a prototypical example being the game `I spy with my
little eyes', as implemented in~\cite{ros2010which}). Perspective taking is a
cognitive skill that relies on the ability for an agent to project itself in
others' `eyes' to estimate what they see from their perspective. While
perspective taking has been previously implemented in robotics by temporary placing
virtual cameras at eyes position for each of the humans tracked by the
robot~\cite{ros2010solving}. a the face of  swapping the different viewpoints
to compute visibility is not directly possible

Geometric \emph{pre-supposition accommodation} makes another interesting case
for alternative worlds representation. \emph{Pre-supposition accommodation}
comes originally from linguistics, where it describes the mechanism by which
\emph{context is adjusted [...] to accept [...] a sentence that imposes certain
requirements on the context in which it is
processed}~\cite{vonfintel2008presupposition}. In the context of spatio-temporal
representations, we call pre-supposition accommodation the ability for an agent
to adjust its model so that it matches some external contextual constraint. For
instance, if A tells B to ``catch the red balloon behind you'', B adds a
representation of a red balloon behind him, even without actually seeing the
balloon: B accommodates the pre-supposition of a red balloon being present
behind himself. Endowing robots with this capability has been touched by
Mavridis \etal within their multi-modal \emph{Grounded Situation
Model}~\cite{Mavridis2006}. However, to the best of our knowledge, a general
framework which would enable robots to derive imaginary worlds by extending of
modifying existing ones has not been proposed so far.
